---
title: "Phase III: Distributional Imputation (Preliminary)"
author: "Deniz Korkmaz"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(tidyverse)
library(zoo)
```

# Data Preparation

Load the cleaned survey dataset and prepare variables for regression-based imputation.

```{r data-preparation}
# Load cleaned data
all_data <- readRDS("cleaned_survey_data.rds")

# Convert relevant columns to numeric and construct derived variables
regression_data <- all_data %>%
  mutate(
    Numerical_Infl_Exp_YR  = as.numeric(as.character(Numerical_Infl_Exp_YR)),
    Categorical_Infl_Exp_YR = as.numeric(as.character(Categorical_Infl_Exp_YR)),
    age                     = as.numeric(as.character(age)),
    time_decimal            = year + (quarter - 1) / 4,
    time_factor             = factor(year + (quarter - 1) / 4),
    age_group               = case_when(
      age < 40              ~ "Young",
      age >= 40 & age <= 60 ~ "Middle",
      age > 60              ~ "Old",
      TRUE                  ~ NA_character_
    )
  ) %>%
  filter(!is.na(age_group))
```

# Cohort Aggregation

Aggregate individual-level responses to cohort-level shares of directional expectations
and mean numerical expectations per time period and age group.

```{r cohort-aggregation}
cohort_data <- regression_data %>%
  group_by(time_decimal, time_factor, age_group) %>%
  summarise(
    p_up      = mean(Categorical_Infl_Exp_YR %in% c(1, 2), na.rm = TRUE),
    p_down    = mean(Categorical_Infl_Exp_YR == 5,          na.rm = TRUE),
    pi_measured = mean(Numerical_Infl_Exp_YR,               na.rm = TRUE),
    .groups   = "drop"
  )
```

# Regression-Based Imputation

Estimate a linear model of mean numerical expectations on directional response shares
and time fixed effects. Apply estimated coefficients to impute expectations for all
cohort-period cells.

```{r imputation}
# Restrict training set to periods with observed numerical expectations
training_data <- cohort_data %>%
  filter(!is.na(pi_measured)) %>%
  mutate(time_factor = droplevels(time_factor))

n_training <- nrow(training_data)
message("Training set size: ", n_training, " cohort-period observations")

if (n_training > 10) {

  model <- lm(pi_measured ~ p_up + p_down + time_factor, data = training_data)

  b_intercept <- coef(model)["(Intercept)"]
  b_up        <- coef(model)["p_up"]
  b_down      <- coef(model)["p_down"]

  cohort_data <- cohort_data %>%
    mutate(
      imputed_inflation = b_intercept + b_up * p_up + b_down * p_down
    )

  message("Imputation complete.")
  head(cohort_data)

} else {
  stop("Insufficient training data for imputation. Check input dataset.")
}
```

# Visualisation

Plot 4-quarter moving-average imputed inflation expectations by age cohort,
following the style of Malmendier & Nagel (2016, Figure 1).

```{r visualisation, fig.width=10, fig.height=6}
plot_data <- cohort_data %>%
  filter(!is.na(age_group)) %>%
  arrange(age_group, time_decimal) %>%
  group_by(age_group) %>%
  mutate(
    smoothed_inflation = rollmean(imputed_inflation, k = 4, fill = NA, align = "right")
  ) %>%
  ungroup()

ggplot(plot_data, aes(x = time_decimal, y = smoothed_inflation, color = age_group)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(
    values = c("Young" = "#E41A1C", "Middle" = "#377EB8", "Old" = "#4DAF4A")
  ) +
  scale_x_continuous(breaks = seq(1950, 2030, 10), expand = c(0, 0)) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title    = "Figure 1: Mean Inflation Expectations by Age Group",
    subtitle = "4-quarter moving average of imputed expectations",
    x        = "Year",
    y        = "Expected Inflation (%)",
    color    = "Age Cohort"
  ) +
  theme_bw() +
  theme(
    plot.title      = element_text(face = "bold", size = 14),
    plot.subtitle   = element_text(size = 11, color = "grey40"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```
