---
title: "Inflation Expectations: Methodological Harmonization"
author: "Deniz Korkmaz"
date: "2026-01-14"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(tidyr)
all_data <- readRDS("~/2804_Inflation/JanFeb26_WorkingFolder/Datasets/Full_Survey_Data/Full_Survey_Cleaned.rds")
apply_curtin_truncation <- TRUE
# TRUE = 1982 until 2026 (Modern) | FALSE = 1982 until 1995 (Original Curtin)
use_modern_beta_period <- TRUE
```


# Categorical Harmonization (Pre-1978) for `Categorical_Infl_Exp_YR`
To ensure longitudinal consistency, we harmonize the pre-1978 response categories of `Categorical_Infl_Exp_YR` with the standard post-1978 structure. The early survey waves included two ambiguous categories
"Stay the same or go up" and "Stay the same or go down", which must be mapped to clear directional responses to align with the modern two-stage question format.

*Pre-1978 Response Categories*:
During the next 12 months, do you think that prices in general will go up, or go down, or stay where they are now?

1.  GO UP  
2.  WILL STAY THE SAME OR GO UP 
3.  SAME
4.  WILL STAY THE SAME OR GO DOWN  
5.  GO DOWN  
8.  DON'T KNOW  
9.  NA: R speaks only of hopes or wishes  

*Post-1978 Response Categories*:
During the next 12 months, do you think that prices in
general will go up, or go down, or stay where they are now?

1.	Go up
2.	Go up (at same rate)
3.	Same
5.	Go down
8.	DK
9.	NA

```{r}
all_data <- all_data %>%
  mutate(Categorical_Infl_Exp_YR = case_when(
    year < 1978 & Categorical_Infl_Exp_YR == 2 ~ 1,
    year < 1978 & Categorical_Infl_Exp_YR == 4 ~ 5,
    TRUE ~ as.numeric(Categorical_Infl_Exp_YR)
  ))
```

Since Malmendier and Nagel (2011) do not explicitly state whether they apply specific outlier truncation, we implement the  `apply_curtin_truncation` lever. This serves as a robustness check to evaluate if extreme values significantly distort the sample means. By toggling this lever, we can compare results using raw data against the standardized truncation at $\pm95\%$ as defined by Curtin (1996).
```{r}
# Truncation-Logik 
if (apply_curtin_truncation) {
  all_data <- all_data %>%
    mutate(Numerical_Infl_Exp_YR = pmin(pmax(Numerical_Infl_Exp_YR, -95), 95))
}
```

# Distributional Imputation of Missing Numerical Values
To address missing information in the quantitative price expectations (Numerical_Infl_Exp_YR), we apply a distributional imputation method as proposed by Curtin (1996).

For respondents who provided a direction (Up/Down) but could not specify an amount, the missing values are imputed by stochastic sampling from the pool of observed responses within the same survey period for the provided direction. 
```{r}
all_data <- all_data %>%
  group_by(year, quarter, Categorical_Infl_Exp_YR) %>%
  mutate(Numerical_Infl_Exp_YR = {
    Numerical_vec <- Numerical_Infl_Exp_YR
        if (unique(Categorical_Infl_Exp_YR) %in% c(1, 5)) {
      
      missing_idx <- is.na(Numerical_vec)
      observed_vals <- Numerical_vec[!missing_idx]
      if (any(missing_idx) && length(observed_vals) > 0) {
        Numerical_vec[missing_idx] <- sample(observed_vals, 
                                            size = sum(missing_idx), 
                                            replace = TRUE)
      }
    }
    Numerical_vec
  }) %>%
  ungroup()
```

# Imputation of Expected Price Declines (Pre-1980)
Prior to February 1980, the survey questionnaire did not include a follow-up question regarding the specific magnitude of expected price declines. This was due to the historical rarity of such expectations during the late 1970s (typically 1% to 2% of respondents).

Following the procedure outlined by Curtin (1996), we assign a fixed value of -3% to all respondents who expected price declines (Category 5) for the period preceding February 1980.
```{r}
all_data <- all_data %>%
  mutate(Numerical_Infl_Exp_YR = as.numeric(Numerical_Infl_Exp_YR)) %>% 
  mutate(Numerical_Infl_Exp_YR = case_when(
    (year < 1980 | (year == 1980 & quarter == 1)) & Categorical_Infl_Exp_YR == 5 ~ -3,
    TRUE ~ Numerical_Infl_Exp_YR
  ))
```

# Adjustment for Misinterpreted "Same" Responses (Pre-1982)
Prior to March 1982, the survey question regarding price expectations was susceptible to a specific measurement bias. As identified by Curtin (1996), many respondents interpreted the option "prices stay the same" as meaning that the inflation rate would remain constant (e.g., staying at 5%), rather than expecting zero percent price growth.

To ensure longitudinal consistency and mitigate this downward bias in the early years of the time series, we implement a two-step dynamic correction for all data preceding the questionnaire change in March 1982 according to Curtin (1996):

+ **Estimation of the Error Probability**: We estimate a period-specific error probability within the "Same" response group. This is based on the observed relationship in post-1982 data between respondents who chose the standard "Up" category and those who specifically clarified they meant "Up at the same rate." 

We model this as: $$P(Error)_t = \beta \times PctUp_t$$ 

Where $PctUp_t$ represents the percentage of respondents expecting price increases in that quarter. Our estimated $\beta$ aligns with the historical coefficient of approximately $0.007$ reported by Curtin (1996) but is updated to reflect the full available data range.

+ **Stochastic Reallocation**: For each quarter before March 1982, we calculate the expected number of misinterpreted responses. A corresponding subset of respondents who answered "Same" (Category 3) is randomly selected for reallocation to the "Up" category.

+ **Distributional Imputation**: Since the numerical expectations for these reallocated cases were not directly recorded (due to the initial "Same" response), we assign them values by sampling from the empirical distribution of reported price increases within the same quarter. This preserves the cross-sectional variance and ensures that the imputed values are representative of the inflationary environment of the time.

```{r}
beta_estimation_end_year <- if(use_modern_beta_period) 2026 else 1995

# Prepare the data for regression from the 1982 transition onwards
regression_data <- all_data %>%
  filter(year >= 1982 & year <= beta_estimation_end_year) %>%
  group_by(year, quarter) %>%
  summarize(
    prop_up = mean(Categorical_Infl_Exp_YR == 1, na.rm = TRUE),           # Category 1
    prop_error = mean(Categorical_Infl_Exp_YR == 2, na.rm = TRUE),        # Category 2 (The mistake)
    prop_same = mean(Categorical_Infl_Exp_YR == 3, na.rm = TRUE),         # Category 3 (The base)
    .groups = "drop"
  ) %>%
  mutate(
    # y = Proportion of errors within the combined 'Same' and 'Error' group
    y = prop_error / (prop_error + prop_same),
    # x = Percentage of 'Up' respondents (0 to 100)
    x = prop_up * 100
  )

# Estimate the model through the origin (no intercept)
curtin_replication_model <- lm(y ~ 0 + x, data = regression_data)
dynamic_beta <- as.numeric(coef(curtin_replication_model))

# Display the result in the console
print(paste("Estimated Beta Coefficient:", round(dynamic_beta, 5)))
```

```{r}
all_data <- all_data %>%
  group_by(year, quarter) %>%
  mutate(
    # 1. Alles sicher innerhalb des move_flag Blocks berechnen
    move_flag = {
      f <- rep(FALSE, n())
      
      # Ist das Quartal vor März 1982?
      is_pre82 <- (year[1] < 1982 | (year[1] == 1982 & quarter[1] == 1))
      
      if (is_pre82) {
        # Skalarer Wert für pct_up (genau EINE Zahl pro Quartal)
        pct_up_scalar <- mean(Categorical_Infl_Exp_YR == 1, na.rm = TRUE) * 100
        
        # Finde die Positionen der Same-Sager
        candidates <- which(Categorical_Infl_Exp_YR == 3)
        
        # Berechne n_to_move (genau EINE Zahl)
        n_to_move <- round(length(candidates) * (dynamic_beta * pct_up_scalar))
        
        # Jetzt funktioniert das && perfekt, da n_to_move Länge 1 hat!
        if (length(candidates) > 0 && !is.na(n_to_move) && n_to_move > 0) {
          selected <- sample(candidates, min(n_to_move, length(candidates)))
          f[selected] <- TRUE
        }
      }
      f
    },
    
    # 2. SYNCHRONISIERTE ZUWEISUNG
    Categorical_Infl_Exp_YR = ifelse(move_flag, 2, Categorical_Infl_Exp_YR),
    
    Numerical_Infl_Exp_YR = {
      v <- Numerical_Infl_Exp_YR
      # WICHTIG: Pool aus der Original-Kategorie 1 ziehen
      up_pool <- v[Categorical_Infl_Exp_YR == 1 & !is.na(v)]
      
      if (any(move_flag) && length(up_pool) > 0) {
        v[move_flag] <- sample(up_pool, sum(move_flag), replace = TRUE)
      }
      v
    }
  ) %>%
  ungroup() %>%
  # Wir müssen nur noch move_flag löschen
  select(-move_flag)
```

Following the stochastic reallocation of misinterpreted responses, all remaining observations in the 'Same' category (Category 3) are assigned a numerical value of 0%.
```{r}
all_data <- all_data %>%
  # 1. Alle restlichen 'Same' Antworten BEDINGUNGSLOS auf 0 setzen
  mutate(Numerical_Infl_Exp_YR = ifelse(Categorical_Infl_Exp_YR == 3, 0, Numerical_Infl_Exp_YR))
```

# Data Checkpoint ()
```{r}
checkpoint_results <- all_data %>%
  mutate(period = ifelse(year < 1982 | (year == 1982 & quarter == 1), "Pre-March 1982", "Post-March 1982")) %>%
  group_by(period) %>%
  summarise(
    # 1. Check: Unvollständige Imputationen (Sollte fast 0 sein)
    missing_num_with_direction = sum(Categorical_Infl_Exp_YR %in% c(1, 5) & is.na(Numerical_Infl_Exp_YR)),
    
    # 2. Check: Durchschnittliche Erwartung der "Same"-Sager
    # Pre-1982 muss dieser Wert > 0 sein (wegen Umschichtung), Post-1982 exakt 0.
    avg_infl_same_group = mean(Numerical_Infl_Exp_YR[Categorical_Infl_Exp_YR == 3], na.rm = TRUE),
    
    # 3. Check: Fixe Zuweisung -3% für Down-Sager (Pre-1980)
    # Wir prüfen den Mittelwert der Down-Sager vor 1980
    avg_down_pre1980 = mean(Numerical_Infl_Exp_YR[year < 1980 & Categorical_Infl_Exp_YR == 5], na.rm = TRUE),
    
    n_obs = n(),
    .groups = "drop"
  )

# Anzeige der Ergebnisse
knitr::kable(checkpoint_results, caption = "Validation of Imputation and Reallocation Steps")
```



# Data Export
```{r}
saveRDS(all_data, "~/2804_Inflation/JanFeb26_WorkingFolder/Datasets/Data_Curtin_Adjustments/Full_survey_data_processed_curtin.rds")
```

