---
title: "Inflation Expectations: Methodological Harmonization"
author: "Deniz Korkmaz"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(tidyr)

# Load cleaned survey data
all_data <- readRDS("cleaned_survey_data.rds")

# Robustness levers
# apply_curtin_truncation: TRUE applies +/- 95% truncation following Curtin (1996)
# use_modern_beta_period:  TRUE uses 1982-present for beta estimation; FALSE uses 1982-1995
apply_curtin_truncation <- TRUE
use_modern_beta_period  <- TRUE
```

# Categorical Harmonization (Pre-1978)

To ensure longitudinal consistency, we harmonize the pre-1978 response categories of
`Categorical_Infl_Exp_YR` with the standard post-1978 structure. The early survey waves
included two ambiguous categories — "Stay the same or go up" and "Stay the same or go down" —
which are mapped to clear directional responses to align with the modern two-stage question format.

**Pre-1978 response categories:**

| Code | Label |
|------|-------|
| 1 | Go up |
| 2 | Will stay the same or go up |
| 3 | Same |
| 4 | Will stay the same or go down |
| 5 | Go down |
| 8 | Don't know |
| 9 | NA |

**Post-1978 response categories:**

| Code | Label |
|------|-------|
| 1 | Go up |
| 2 | Go up (at same rate) |
| 3 | Same |
| 5 | Go down |
| 8 | Don't know |
| 9 | NA |

```{r categorical-harmonization}
all_data <- all_data %>%
  mutate(
    Categorical_Infl_Exp_YR = case_when(
      year < 1978 & Categorical_Infl_Exp_YR == 2 ~ 1,  # "same or up" -> "up"
      year < 1978 & Categorical_Infl_Exp_YR == 4 ~ 5,  # "same or down" -> "down"
      TRUE                                        ~ as.numeric(Categorical_Infl_Exp_YR)
    )
  )
```

# Outlier Truncation (Curtin 1996)

Since Malmendier & Nagel (2016) do not explicitly specify outlier treatment, we implement
the `apply_curtin_truncation` lever as a robustness check. When enabled, numerical
expectations are winsorized at ±95%, following Curtin (1996).

```{r outlier-truncation}
if (apply_curtin_truncation) {
  all_data <- all_data %>%
    mutate(
      Numerical_Infl_Exp_YR = pmin(pmax(Numerical_Infl_Exp_YR, -95), 95)
    )
}
```

# Distributional Imputation of Missing Numerical Values

To address missing quantitative price expectations (`Numerical_Infl_Exp_YR`), we apply
the distributional imputation method of Curtin (1996). Respondents who provided a
directional response (Up/Down) but no quantity receive imputed values sampled from the
empirical distribution of observed responses within the same survey period and direction cell.

```{r distributional-imputation}
all_data <- all_data %>%
  group_by(year, quarter, Categorical_Infl_Exp_YR) %>%
  mutate(
    Numerical_Infl_Exp_YR = {
      vals <- Numerical_Infl_Exp_YR
      # Only impute for directional responders (Up = 1/2, Down = 5)
      if (unique(Categorical_Infl_Exp_YR) %in% c(1, 2, 5)) {
        missing_idx   <- is.na(vals)
        observed_vals <- vals[!missing_idx]
        if (any(missing_idx) && length(observed_vals) > 0) {
          vals[missing_idx] <- sample(observed_vals,
                                      size    = sum(missing_idx),
                                      replace = TRUE)
        }
      }
      vals
    }
  ) %>%
  ungroup()
```

# Imputation of Expected Price Declines (Pre-1980)

Prior to February 1980, the survey did not ask respondents for the magnitude of expected
price declines, as such expectations were historically rare (typically 1–2% of respondents).
Following Curtin (1996), we assign a fixed value of **-3%** to all pre-1980 observations
with a directional "down" response (Category 5).

```{r pre-1980-declines}
all_data <- all_data %>%
  mutate(
    Numerical_Infl_Exp_YR = as.numeric(Numerical_Infl_Exp_YR),
    Numerical_Infl_Exp_YR = case_when(
      (year < 1980 | (year == 1980 & quarter == 1)) &
        Categorical_Infl_Exp_YR == 5 ~ -3,
      TRUE                            ~ Numerical_Infl_Exp_YR
    )
  )
```

# Bias Correction for "Same" Responses (Pre-1982)

Prior to March 1982, many respondents interpreted "prices stay the same" as meaning
the *inflation rate* stays constant rather than *zero price growth* (Curtin 1996). We
implement a two-step stochastic correction to mitigate this downward bias.

**Step 1: Estimate the period-specific error probability**

We model the share of misinterpreted "Same" responses as proportional to the share
of "Up" respondents in the same period:

$$P(\text{Error})_t = \beta \times \text{PctUp}_t$$

The coefficient $\beta \approx 0.007$ follows Curtin (1996) but is estimated on the
full available post-1982 data when `use_modern_beta_period = TRUE`.

**Step 2: Stochastic reallocation**

For each pre-1982 quarter, the estimated number of misinterpreted responses is randomly
reallocated from Category 3 ("Same") to Category 1 ("Up"). Imputed numerical values
are drawn from the empirical distribution of observed "Up" responses in the same quarter.

```{r pre-1982-bias-correction}
# Estimate beta from post-1982 data
if (use_modern_beta_period) {
  beta_data <- all_data %>% filter(year >= 1982)
} else {
  beta_data <- all_data %>% filter(year >= 1982, year <= 1995)
}

period_shares <- beta_data %>%
  group_by(year, quarter) %>%
  summarise(
    pct_up   = mean(Categorical_Infl_Exp_YR %in% c(1, 2), na.rm = TRUE),
    pct_same = mean(Categorical_Infl_Exp_YR == 3,          na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  filter(pct_up > 0)

beta_hat <- coef(lm(pct_same ~ 0 + pct_up, data = period_shares))["pct_up"]
message(sprintf("Estimated beta: %.4f (Curtin 1996 reference: ~0.007)", beta_hat))

# Apply stochastic reallocation to pre-1982 data
set.seed(42)

corrected_data <- all_data %>%
  group_by(year, quarter) %>%
  group_modify(function(df, keys) {

    if (keys$year >= 1982 | (keys$year == 1982 & keys$quarter >= 1)) {
      return(df)
    }

    pct_up <- mean(df$Categorical_Infl_Exp_YR %in% c(1, 2), na.rm = TRUE)
    n_reallocate <- round(beta_hat * pct_up * nrow(df))

    same_idx <- which(df$Categorical_Infl_Exp_YR == 3)
    up_vals  <- df$Numerical_Infl_Exp_YR[df$Categorical_Infl_Exp_YR %in% c(1, 2) &
                                           !is.na(df$Numerical_Infl_Exp_YR)]

    if (n_reallocate > 0 && length(same_idx) >= n_reallocate && length(up_vals) > 0) {
      selected_idx <- sample(same_idx, size = n_reallocate, replace = FALSE)
      df$Categorical_Infl_Exp_YR[selected_idx]  <- 1
      df$Numerical_Infl_Exp_YR[selected_idx]    <- sample(up_vals,
                                                           size    = n_reallocate,
                                                           replace = TRUE)
    }

    df
  }) %>%
  ungroup()

all_data <- corrected_data
message("Bias correction complete.")
```
