---
title: "Inflation Expectations: Methodological Harmonization"
author: "Deniz Korkmaz"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(ggplot2)
library(tidyr)

all_data <- readRDS("cleaned_survey_data.rds")

# toggle these for robustness checks
apply_curtin_truncation <- TRUE
use_modern_beta_period  <- TRUE
```

# Categorical Harmonization (Pre-1978)

Pre-1978 surveys used a slightly different response scale. Categories 2 ("same or up")
and 4 ("same or down") don't exist post-1978, so we recode them to the nearest
directional equivalent before doing anything else.

```{r}
all_data <- all_data %>%
  mutate(
    Categorical_Infl_Exp_YR = case_when(
      year < 1978 & Categorical_Infl_Exp_YR == 2 ~ 1,
      year < 1978 & Categorical_Infl_Exp_YR == 4 ~ 5,
      TRUE ~ as.numeric(Categorical_Infl_Exp_YR)
    )
  )
```

# Outlier Truncation

M&N (2016) don't explicitly mention truncation, so this is implemented as an optional
robustness check via the `apply_curtin_truncation` flag above. Curtin (1996) suggests
capping at ±95%.

```{r}
if (apply_curtin_truncation) {
  all_data <- all_data %>%
    mutate(Numerical_Infl_Exp_YR = pmin(pmax(Numerical_Infl_Exp_YR, -95), 95))
}
```

# Distributional Imputation of Missing Numerical Values

Following Curtin (1996): respondents who gave a direction (up/down) but no number
get imputed by sampling from others in the same period + direction cell.

```{r}
all_data <- all_data %>%
  group_by(year, quarter, Categorical_Infl_Exp_YR) %>%
  mutate(
    Numerical_Infl_Exp_YR = {
      vals <- Numerical_Infl_Exp_YR
      if (unique(Categorical_Infl_Exp_YR) %in% c(1, 2, 5)) {
        miss <- is.na(vals)
        obs  <- vals[!miss]
        if (any(miss) && length(obs) > 0)
          vals[miss] <- sample(obs, size = sum(miss), replace = TRUE)
      }
      vals
    }
  ) %>%
  ungroup()
```

# Pre-1980: Fixed Value for Expected Declines

Before Feb 1980 the survey didn't ask for a magnitude when someone expected prices
to go down (barely anyone did back then). We follow Curtin (1996) and assign -3%.

```{r}
all_data <- all_data %>%
  mutate(
    Numerical_Infl_Exp_YR = as.numeric(Numerical_Infl_Exp_YR),
    Numerical_Infl_Exp_YR = case_when(
      (year < 1980 | (year == 1980 & quarter == 1)) & Categorical_Infl_Exp_YR == 5 ~ -3,
      TRUE ~ Numerical_Infl_Exp_YR
    )
  )
```

# Pre-1982: "Same" Response Bias Correction

Curtin (1996) documents that before the March 1982 questionnaire change, many
respondents interpreted "same" as "inflation rate stays the same" rather than
"zero price growth". We correct for this stochastically:

1. Estimate $\beta$ from post-1982 data: $P(\text{error})_t \approx \beta \times \text{PctUp}_t$
2. For each pre-1982 quarter, randomly reallocate that many "Same" responses to "Up"
   and impute numerical values from the observed "Up" distribution in that quarter.

```{r}
beta_data <- if (use_modern_beta_period) {
  all_data %>% filter(year >= 1982)
} else {
  all_data %>% filter(year >= 1982, year <= 1995)
}

period_shares <- beta_data %>%
  group_by(year, quarter) %>%
  summarise(
    pct_up   = mean(Categorical_Infl_Exp_YR %in% c(1, 2), na.rm = TRUE),
    pct_same = mean(Categorical_Infl_Exp_YR == 3, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  filter(pct_up > 0)

beta_hat <- coef(lm(pct_same ~ 0 + pct_up, data = period_shares))["pct_up"]

set.seed(42)

all_data <- all_data %>%
  group_by(year, quarter) %>%
  group_modify(function(df, keys) {

    if (keys$year >= 1982) return(df)

    pct_up       <- mean(df$Categorical_Infl_Exp_YR %in% c(1, 2), na.rm = TRUE)
    n_reallocate <- round(beta_hat * pct_up * nrow(df))
    same_idx     <- which(df$Categorical_Infl_Exp_YR == 3)
    up_vals      <- df$Numerical_Infl_Exp_YR[
      df$Categorical_Infl_Exp_YR %in% c(1, 2) & !is.na(df$Numerical_Infl_Exp_YR)
    ]

    if (n_reallocate > 0 && length(same_idx) >= n_reallocate && length(up_vals) > 0) {
      idx <- sample(same_idx, n_reallocate, replace = FALSE)
      df$Categorical_Infl_Exp_YR[idx] <- 1
      df$Numerical_Infl_Exp_YR[idx]   <- sample(up_vals, n_reallocate, replace = TRUE)
    }

    df
  }) %>%
  ungroup()
```
